---
title: "Reuse and correct accepted research"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reuse and correct accepted research}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A synthetic pilot first reports 82% material recovery. A later correction says
62%. The host accepts each completed research product, reviews which claim is
superseded, and retains the exact evidence used for the first answer.

This example runs without credentials or network access. Its two saved promotion
bundles were produced from completed Tempest STORM fixtures using
`tempest_promotion_bundle()` and `tempest_save_promotion_bundle()`. Their reports
come from `tempest_report()`. The fixture builder is
`tools/rebuild-accepted-research.R` in the source repository. These are synthetic
inputs for exercising acceptance and reuse, not measured pilot results or a
claim that a live model researched this topic.

```{r run}
library(tempest)
example <- new.env()
sys.source(system.file("examples", "accepted-research.R", package = "tempest"), example)
result <- example$accepted_research_example()
result$selected_records
result$reviewed_changes[, c("class", "action")]
result$original_preserved
```

The example creates and closes a temporary durable Graft store. It reopens the
store and checkpoint before an unchanged day. It then accepts the correction
and separately commits the host's decision to supersede the older claim.
This is an explicit review decision; differing model outputs alone do not
invalidate accepted knowledge.

Each selected bundle retains its claim, support, evidence span and source.
Multiple selections form a union; an unchanged day preserves the full union.
The recipe stops if that complete evidence exceeds 1,000 records, rather than
keeping claims while discarding their support. Selected-record changes,
including deletion tombstones, require review before another automatic run.

The earlier checkpoint still reconstructs the earlier accepted revisions:

```{r reports, results='asis'}
cat(result$initial_report, "\n\n", result$corrected_report)
```

The reports retain their execution-review annotations, including unknown stage
support and publication blocks. Host acceptance of the selected evidence does
not clear those diagnostics or establish publishable research quality.

Readable synthesis remains host-owned checkpoint content. Graft stores the
accepted research evidence through Tempest's existing schema; this example
adds no report class or executable procedure. The checkpoint links a report to
its exact selection and accepted boundary, but does not prove every sentence.
Source identity, source-content hash and captured excerpts remain distinct from
that prose. File retention, sharing, atomic checkpoint publication and erasure
are responsibilities of the application.

The [daily briefing](daily-briefing.html) applies the same selection recipe to a
scheduled host. Its research step is optional and may use a live model; the
acceptance/restart/correction example above is entirely offline.
