---
title: "Use Tempest Agent Skills"
description: >
  Install Tempest's two research Agent Skills or make portable STORM research
  callable from an ellmer and shinychat application.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use Tempest Agent Skills}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Tempest ships Agent Skills alongside its R APIs. Two research skills are
supported. Each is a portable directory containing a `SKILL.md` file,
supporting references, and agent metadata. A compatible agent can load those
instructions when it needs to operate Tempest or conduct STORM research in
another framework.

This article covers two deployment shapes:

1. install the skills for a coding or research agent; or
2. let an ellmer client load a skill at runtime inside shinychat.

The second shape makes a skill callable, but the skill remains instructions.
The host application must still provide the tools, authorization, state, and
lifecycle needed to perform the research.

## Agent Skills and product APIs are different

An Agent Skill is external operating guidance installed with
`tempest_agent_skills()` or `tempest_install_agent_skills()`. It does not grant
runtime authority. The host still provides the tools, authorization, state,
and lifecycle needed to perform the research.

## Choose a supported research skill

Use the smallest skill that matches the task:

| Skill | Lifecycle | Use it when the agent should |
|---|---|---|
| `use-tempest-research` | Supported | Choose, configure, run, resume, inspect, or embed Tempest's built-in STORM and Co-STORM workflows |
| `conduct-storm-research` | Supported | Follow a provider- and framework-neutral STORM or Co-STORM protocol outside Tempest |

The two research skills describe complete operating modes and can be used
independently.

## Discover and install the skills

List the skill directories bundled with the installed package:

```{r discover}
library(tempest)

skills <- tempest_agent_skills()
names(skills)
```

Install the supported skills into a directory used by the agent:

```{r install-selected, eval = FALSE}
tempest_install_agent_skills(
  "~/.codex/skills",
  skills = c("use-tempest-research", "conduct-storm-research")
)
```

Existing skill directories are preserved by default. Set `overwrite = TRUE`
only when the installed copy should be replaced by the version bundled with
the current Tempest package.

After installation, ask the agent to load the relevant skill by name. A useful
request names the desired outcome rather than restating the procedure:

> Load `use-tempest-research` and help me configure a resumable STORM report.

## Let an ellmer client discover Tempest skills

The [btw skill
tool](https://posit-dev.github.io/btw/reference/btw_tool_skill.html) discovers
skills bundled in attached R packages. Attach Tempest before registering the
`"skills"` tool group so the resolved skill paths include `tempest`:

```{r skill-loader, eval = FALSE}
library(tempest)
library(ellmer)
library(btw)

client <- ellmer::chat(
  "openai/gpt-5.6-sol",
  system_prompt = paste(
    "Load conduct-storm-research for deep research requests.",
    "Follow the loaded protocol and use only the registered research tools."
  )
)

client$register_tools(btw::btw_tools("skills"))
```

Use a `btw` release that provides the `"skills"` tool group. Skills are
resolved when the tools are registered, so create a new client after changing
skill paths or attached packages.

Registering only the skill tool lets the model read the instructions and
references. It does not provide search, evidence storage, report publication,
or persistent Co-STORM state.

## Make portable research callable in shinychat

`conduct-storm-research` is the right boundary when a shinychat application
should offer STORM research without depending on Tempest's workflow APIs. The
outer ellmer client loads the protocol, then uses narrow tools supplied by the
host application.

At minimum, the host tool set should cover:

| Capability | Batch STORM | Co-STORM |
|---|---:|---:|
| Search approved sources | Required | Required |
| Read full source content | Required | Required |
| Save stable source and claim records | Required | Required |
| Publish a report artifact | Required | Required |
| Persist checkpoints | Recommended | Required |
| Preserve expert, transcript, and mind-map state | No | Required |
| Report progress and cancel work | Recommended | Recommended |

Do not substitute search snippets for full-source reading or the model's
conversation window for durable Co-STORM state.

### Single-user prototype

For a local, single-user prototype, attach Tempest, create an ellmer client,
register the skill loader and host tools, and pass the client to
`chat_app()`:

```{r shinychat-prototype, eval = FALSE}
library(tempest)
library(ellmer)
library(btw)
library(shinychat)

client <- ellmer::chat(
  "openai/gpt-5.6-sol",
  system_prompt = paste(
    "A portable STORM skill is available.",
    "Load conduct-storm-research for deep research requests.",
    "Default to batch mode unless the user asks for collaborative research.",
    "Use only the registered research and evidence tools."
  )
)

host_tools <- research_tools_for_session()

client$register_tools(c(
  btw::btw_tools("skills"),
  host_tools
))

shinychat::chat_app(client)
```

`research_tools_for_session()` is an application-owned constructor, not a
Tempest function. It should return fresh `ellmer::tool()` definitions for
approved search, full-source reading, evidence writes, artifact publication,
progress, cancellation, and any persistent state required by the selected
mode. Keep credentials and authenticated clients inside the tool closures;
never return them to the model or save them in a checkpoint.

The [`btw_tools()`
reference](https://posit-dev.github.io/btw/reference/btw_tools.html) describes
tool-group registration. The [ellmer tool-calling
guide](https://ellmer.tidyverse.org/articles/tool-calling.html) covers typed
arguments, tool results, and the automatic tool loop.

### Multi-user shinychat application

`chat_app()` mutates one client and is intended for a single user. In a
multi-user Shiny application, create the ellmer client and host tool state
inside the server function, then use `chat_mod_server()`:

```{r shinychat-multi-user, eval = FALSE}
library(tempest)
library(shiny)
library(bslib)
library(ellmer)
library(btw)
library(shinychat)

ui <- bslib::page_fillable(
  bslib::card(
    full_screen = TRUE,
    shinychat::chat_mod_ui("research")
  )
)

server <- function(input, output, session) {
  client <- ellmer::chat(
    "openai/gpt-5.6-sol",
    system_prompt = paste(
      "Load conduct-storm-research for deep research requests.",
      "Use only the registered research and evidence tools."
    )
  )

  client$register_tools(c(
    btw::btw_tools("skills"),
    research_tools_for_session(session)
  ))

  shinychat::chat_mod_server(
    "research",
    client,
    greeting = paste(
      "Ask me for a multi-perspective research report.",
      "I will show tool calls and preserve cited evidence."
    )
  )
}

shiny::shinyApp(ui, server)
```

The [shinychat module
reference](https://posit-dev.github.io/shinychat/r/reference/chat_app.html)
documents the single-user and multi-user distinction. The per-session
constructor is essential: sharing the mutable ellmer client, evidence ledger,
or tool closures across sessions can mix users' conversations and data.

Run long research outside the reactive main path, connect the UI's stop action
to the active research boundary, and keep safe partial evidence when work is
cancelled. Return a concise tool result to the transcript and expose the full
report, sources, claims, and checkpoint through appropriate Shiny UI or
downloads.

## Choose dynamic loading or one research tool

Dynamic loading gives the outer model access to the skill and the underlying
research tools. It is flexible and transparent, but the outer model must
select and follow the protocol.

For a tighter application boundary, wrap the protocol and tools in a dedicated
research agent, then register that agent as one outer tool such as
`research_topic(topic, mode, constraints)`. This shape is useful when the host
needs fixed budgets, simpler authorization, or a compact tool result. The
outer chat should receive a structured status, report, evidence summary, and
checkpoint identifier rather than only prose.

Whichever shape you choose, keep the research state separate from the outer
chat history.

## Verify the integration

Use deterministic local search and evidence tools before enabling live
providers. Exercise:

- discovery and loading of `conduct-storm-research`;
- a complete batch research request;
- refusal of Co-STORM when persistent state is unavailable;
- missing, weak, and conflicting evidence;
- per-session isolation;
- cancellation and partial checkpoint behavior;
- rich tool-call display and a plain-text fallback; and
- absence of credentials or live clients in saved state.

The integration is ready when the same evidence and artifact contracts survive
success, failure, cancellation, and restore.
