---
title: "Reuse retained artifact evidence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reuse retained artifact evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A research run can use an exact set of retained evidence without opening a Graft
store. `tempest_artifact_knowledge()` accepts a host-selected description and its
text content. The host owns storage, saving, review, purpose, access, and current
consultation eligibility. Tempest owns the research interpretation of that input.

## Build an exact input

This small example represents two artifacts already retained and selected by a
host. In an application, read the references and content from the artifact store
and check current eligibility before constructing the input.

```{r}
library(tempest)
contents <- list(
  "claim:pilot" = "statement_text: The pilot recovered 82%.\nstatus: active",
  "source:pilot" = "excerpt: The pilot recovered 82%."
)
refs <- lapply(names(contents), function(id) {
  list(
    record_id = id,
    revision_id = "revision:1",
    class = if (startsWith(id, "claim")) "Claim" else "Source",
    sha256 = digest::digest(charToRaw(enc2utf8(contents[[id]])),
      algo = "sha256", serialize = FALSE),
    dependencies = if (startsWith(id, "claim")) {
      list(list(record_id = "source:pilot", revision_id = "revision:1"))
    } else {
      list()
    }
  )
})
selection <- list(
  selection_id = "selection:pilot-briefing-1",
  purpose = "daily-briefing",
  records = refs,
  provenance = list(producer = "synthetic-pilot")
)
knowledge <- tempest_artifact_knowledge(selection, contents)
knowledge
```

Each selected record has an immutable revision reference and the SHA-256 digest
of its UTF-8 content bytes. The contents must cover every selected record exactly
once. Every declared dependency must resolve to the exact selected revision.
Supported evidence classes are Claim, ClaimSupport, EvidenceSpan, and Source.
The host supplies the complete required selection; Tempest does not infer omitted
relationships or authenticate the host's acceptance decision.

Only inert text is admitted. Claims use the accepted projection's
`statement_text: ...` and `status: ...` fields for briefing change detection.
An active statement already present in this input can support a no-change item;
a superseded or retracted statement cannot. Storing code or instructions does not
make them tools, prompts, or governed procedures.

## Use it in a new run

```{r, eval=FALSE}
result <- tempest_run("Pilot recovery briefing", knowledge = knowledge)
session <- tempest_session("Pilot recovery briefing", knowledge = knowledge)
tempest_session_save(session, "briefing-session")

# Check the host's current access and eligibility policy again before resuming.
resumed <- tempest_session_resume("briefing-session", config = tempest_config())
```

The workspace retains the full input selection and evidence resources. Session
bundles include the selection in `session.json` and checksummed resource content.
STORM bundles retain it in their workspace identity and evidence snapshot. The
selection's input digest binds its description, separately from a native Graft
snapshot, receipt, or acceptance event. Source receipts and other source
identities can be retained as inert `provenance` data.

A correction gets a new artifact revision and a new selection. A new run can use
that selection while the previous run still resolves its original evidence.
Saving and resuming preserves the historical input; it does not freeze permission
or automatically reapprove withdrawn knowledge. The host must enforce current
eligibility before starting or resuming work. This constructor does not monitor
policy changes during an active run.

## Scope and current formats

The input is bounded to 1,000 records, 1 MiB of selection metadata, and 1 MiB of
text content. This slice supports evidence admission and cross-run reuse. Future
acceptance into an artifact store, concurrent publication, permanent deletion,
and production recovery remain host/storage work. Existing Graft promotion is
still available for the reference producer; its presence is not a decision to
retain Graft.

Workspace format 6, STORM format 9, and Co-STORM format 12 retain artifact inputs.
This project is pre-production: regenerate older experimental bundles instead of
using a compatibility reader.
