---
title: "Build a governed daily briefing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Build a governed daily briefing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

A useful daily briefing is an attention system: it says what changed, why the
change matters, what deserves review today, and what important topics did not
change. It should not turn every observation into accepted organizational
knowledge.

This workflow needs no briefing-specific Tempest interface. The host owns the
schedule, monitored topic, presentation, and approval policy. The packages keep
their existing responsibilities:

| Module | Responsibility in the briefing |
|---|---|
| Graft | Pin accepted knowledge before research, report what was accepted since the last run, and hold reviewed revisions afterward. |
| Tempest | Produce a source-grounded research product whose change signal is computed against the pinned snapshot, plus a review-only evidence proposal. |
| dsprrr | Run Tempest's validated structured research stages. |
| Deputy | Execute Tempest's permission-bounded agent work. |
| scans | Inspect the closed Tempest trajectory without reopening live product state. |

The host directly calls Tempest, Graft, and scans. Tempest uses dsprrr and
Deputy behind its product interface, so the host does not need to coordinate
their internal objects.

## Pin what the organization already knows

Open one Graft store with Tempest's schema. Capture an immutable view before
starting the run, then select the accepted evidence that is relevant to this
briefing. An empty selection is valid on the first day.

```{r setup}
library(tempest)
library(graft)
library(scans)

store <- graft_open(
  tempest_graft_schema(),
  "briefing/knowledge.duckdb",
  okf = "managed"
)

topic <- paste(
  "Daily decision briefing on grid-scale battery recycling as of",
  Sys.Date(),
  "Cover what changed, why it matters, what needs review today,",
  "and any material no-change signal explicitly established by evidence.",
  "Omit no-change when evidence is missing or inconclusive. Preserve",
  "uncertainty and cite evidence."
)
```

The selection recipe below pins the accepted boundary for the whole run. A
commit made by another process cannot silently alter its input.

### Select the accepted basis durably

The host selects complete promotion bundles, including each claim's supports,
evidence spans and sources. It keeps the union of those selections across
acceptance cycles. An unchanged day retains the same checkpoint; it does not
replace the evidence with an empty change receipt or fill a record limit by
dropping older sources.

For repeated record IDs, the newest selected receipt revision wins. Capturing a
checkpoint stops if current evidence differs from those reviewed revisions, so
a concurrent update cannot silently become part of an accepted report's basis.

This example uses the small host-owned recipe shipped with Tempest. It is an
example, not a new package API or an access-control mechanism. The checkpoint
is trusted local host data; authorize access and consultation before reading it.

```{r basis}
source(system.file("examples", "briefing-basis.R", package = "tempest"))
basis_path <- "briefing/basis.rds"
if (file.exists(basis_path)) {
  basis <- readRDS(basis_path)
  changes <- briefing_changes(store, basis)
  if (nrow(changes) > 0L) {
    stop("Selected evidence changed; review the complete selection before research.")
  }
} else {
  basis <- capture_briefing_basis(store, selections = list())
}
knowledge <- read_briefing_basis(store, basis)
```

`briefing_changes()` uses `graft_changes(record_ids = ...)` to retain changes to
selected records, including deletion tombstones whose payload no longer names
their parent claim. An unrelated acceptance does not join the selection. A
changed or superseded claim pauses automatic reuse until the host reviews the
selection. Historical inspection can still use the original checkpoint.

The complete selection must fit `tempest_knowledge()`'s 1,000-record ceiling.
Exceeding it stops the example rather than silently losing evidence. A snapshot
fixes an accepted boundary; the checkpoint additionally retains the selected
IDs, exact revisions and host-accepted report text. The report is synthesis,
not primary source evidence, and its recorded basis does not establish support
for every sentence.

## Research today's change

Use a small expert panel that reflects the decision, not a collection of
generic personas. The topic asks for the briefing's decision structure while
Tempest keeps the exact report, sources, claims, and claim-support pairs behind
its ordinary product interface.

```{r research}
technical <- tempest_expert(
  name = "Technical readiness reviewer",
  title = "Battery recycling process specialist",
  description = "Reviews demonstrated process performance and scale-up risk.",
  instructions = paste(
    "Separate measured performance from projections.",
    "Call out consequential unknowns. Report no change only when a verified",
    "claim restates a condition the organization already accepted."
  ),
  focus_areas = c("process yield", "scale-up", "safety")
)

market <- tempest_expert(
  name = "Market and policy reviewer",
  title = "Battery supply-chain analyst",
  description = "Reviews policy, capacity, partnerships, and market movement.",
  instructions = paste(
    "Distinguish announcements from operating evidence.",
    "State which developments could change a decision today."
  ),
  focus_areas = c("policy", "capacity", "partnerships", "economics")
)

result <- tempest_run(
  topic,
  config = tempest_config(
    models = "openai/gpt-5.6-luna",
    citation_policy = "claim_verified",
    on_unsupported_claim = "flag"
  ),
  knowledge = knowledge,
  experts = list(technical, market),
  max_questions_per_perspective = 2,
  output_dir = "briefing/runs",
  run_id = paste0("daily-", Sys.Date())
)

report <- tempest_report(result)
sources <- tempest_sources(result)
claims <- tempest_claims(result)
supports <- tempest_claim_supports(result)
```

Treat the report as the briefing and the tables as its evidence rail. The
at-a-glance brief and each detailed section keep four statements visibly
distinct, and the first and last of them are decided against the pinned
snapshot rather than by the prompt:

- "What changed" items copy verified claims exactly, and Tempest accepts one
  only when its text does not restate a `Claim` already in the pinned
  knowledge. On the first day every verified claim is new.
- "Why it matters" assessments are labeled, confidence-calibrated inferences.
- "Review today" items are proposed actions.
- "No material change" statements copy one verified claim exactly, and
  Tempest accepts one only when that claim restates an accepted `Claim`. Today's
  evidence re-verified something the organization already knew.

Missing, unresolved, or inconclusive evidence does not become a no-change
statement, because it does not produce a verified claim at all. Tempest binds
every item to its exact claim IDs in machine-readable provenance, re-checks
each item's disposition against the accepted claims whenever the report is
rendered, and rejects a report if that provenance no longer matches the
visible Markdown.

Whether somebody was shown the briefing or read it is operational state; it is
not accepted knowledge. An assessment or review action is also not accepted
knowledge merely because it appears in the briefing.

## Review the evidence and the run separately

Prepare a Graft plan, but do not commit it. In parallel, project the completed
run into scans. The first review asks whether the proposed knowledge is fit to
accept; the second asks whether the agent trajectory itself behaved well.

```{r review}
proposal <- tempest_promotion_bundle(result)
plan <- tempest_graft_plan(store, proposal)

review <- tempest_trajectory_review(
  result,
  promotion_bundle = proposal
)
trajectory <- scans::as_trajectory_tempest(review)

run_summary <- scans::summarize_trajectories(trajectory)
run_findings <- scans::scan_trajectories(trajectory)

plan@valid
plan@changes[, c("class", "record_id", "action", "disposition", "changed_fields")]
run_summary
run_findings
```

Tempest keys accepted `Claim` identity on the claim's normalized statement
text and accepted `Source` identity on its locator plus content hash. A claim the organization
already accepted therefore plans as an `update` whose `changed_fields` name
only run metadata such as the Tempest claim id and verification time, and
Graft's `disposition` column reads `revision`; a genuinely new statement reads
`new`; an identical proposal reads `duplicate`. A staged claim that sets
`superseded_by`, or evidence whose `support_type` is `contradicts`, surfaces
as `superseded`, `supersedes`, or `contradicted` so the reviewer sees
statement-level relations without reading every row.

Do not collapse these into one confidence score. Claim support, proposed Graft
changes, and trajectory findings answer different review questions.

## Accept deliberately, including no change

Accepted knowledge changes only when a `Claim` is new, superseded, or
contradicted. A plan whose claims are all `duplicate` or `revision` proposes
nothing the organization does not already know; publish the briefing and
record that no-change outcome in the host's run log rather than manufacturing a
knowledge revision to prove the schedule ran.

When the reviewer accepts a real change, commit the exact plan that was shown
and bind a Tempest receipt to the resulting immutable snapshot. Save the basis
the next run will carry forward either way, so a first day without accepted
change still pins its selection instead of searching again tomorrow:

```{r accept}
claim_rows <- plan@changes[plan@changes$class == "Claim", ]
has_change <- any(
  claim_rows$disposition %in% c("new", "supersedes", "superseded", "contradicted")
)
approved <- FALSE # The host supplies the review decision.

if (has_change && approved) {
  commit_result <- graft_commit(store, plan)
  receipt <- tempest_promotion_receipt(store, proposal, plan, commit_result)
  accepted_review <- tempest_trajectory_review(
    result, promotion_bundle = proposal, promotion_receipt = receipt
  )
  selections <- c(basis$selections, list(briefing_selection(receipt)))
  basis <- capture_briefing_basis(store, selections, tempest_report(result))
}
saveRDS(basis, basis_path)
```

A rejected or unchanged proposal preserves the previous basis and accepted
report. Today's unaccepted briefing belongs in the host's run log. A correction
requires reviewing the older claim's lifecycle and replacing its selection
with the complete corrected bundle; appending the new claim alone does not
supersede the old one. Keep the prior checkpoint for authorized historical
inspection, subject to the host's erasure policy.

Commit the reviewed plan before saving its checkpoint. If saving fails, report
the checkpoint failure separately from acceptance. Recover from the retained
plan, bundle, commit result and receipt; do not report a failed commit or retry
with a fabricated receipt. This short recipe uses one local host writer; a
production scheduler needs its own atomic checkpoint publication and recovery.

The offline [accepted-research example](accepted-research.html) demonstrates
two contradictory research products, an unchanged day and explicit correction
without API keys.

## Let a person edit accepted knowledge

The store was opened with `okf = "managed"`, so Graft keeps a readable Open
Knowledge Format working tree next to the DuckDB file. A reviewer who wants to
correct a wording, retract a claim, or mark one statement as superseding
another edits those files; Graft turns the edits back into the same kind of
plan the briefing produced, and the same acceptance step applies.

```{r okf}
graft_sync(store)
graft_status(store)$status
#> "current"

# A person edits files under the managed OKF directory, then:
graft_status(store)$status
#> "modified"

edited <- graft_review(
  store,
  provenance = graft_provenance(
    "briefing-reviewer",
    idempotency_key = paste0("okf-", Sys.Date())
  )
)
edited@changes[, c("class", "record_id", "action", "disposition")]

if (edited@valid) {
  graft_commit(store, edited)
  graft_sync(store)
}
```

The edited plan is bound to the exact bundle and accepted batch it was
reviewed against, so a commit that landed in between fails the plan instead
of silently rebasing the person's edits. Tomorrow's briefing picks the edit up
through `graft_changes()` like any other accepted change.

```{r close}
graft_close(store)
```
