---
title: "Build a governed daily briefing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Build a governed daily briefing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

A useful daily briefing is an attention system: it says what changed, why the
change matters, what deserves review today, and what important topics did not
change. It should not turn every observation into accepted organizational
knowledge.

This workflow needs no briefing-specific Tempest interface. The host owns the
schedule, monitored topic, presentation, and approval policy. The packages keep
their existing responsibilities:

| Module | Responsibility in the briefing |
|---|---|
| Graft | Pin accepted knowledge before research, report what was accepted since the last run, and hold reviewed revisions afterward. |
| Tempest | Produce a source-grounded research product whose change signal is computed against the pinned snapshot, plus a review-only evidence proposal. |
| dsprrr | Run Tempest's validated structured research stages. |
| Deputy | Execute Tempest's permission-bounded agent work. |
| scans | Inspect the closed Tempest trajectory without reopening live product state. |

The host directly calls Tempest, Graft, and scans. Tempest uses dsprrr and
Deputy behind its product interface, so the host does not need to coordinate
their internal objects.

## Pin what the organization already knows

Open one Graft store with Tempest's schema. Capture an immutable view before
starting the run, then select the accepted evidence that is relevant to this
briefing. An empty selection is valid on the first day.

```{r setup}
library(tempest)
library(graft)
library(scans)

store <- graft_open(
  tempest_graft_schema(),
  "briefing/knowledge.duckdb",
  okf = "managed"
)

snapshot <- graft_snapshot(store)
view <- graft_at(store, snapshot)

topic <- paste(
  "Daily decision briefing on grid-scale battery recycling as of",
  Sys.Date(),
  "Cover what changed, why it matters, what needs review today,",
  "and any material no-change signal explicitly established by evidence.",
  "Omit no-change when evidence is missing or inconclusive. Preserve",
  "uncertainty and cite evidence."
)
```

The snapshot fixes the accepted boundary for the whole run. A commit made by
another process halfway through the briefing cannot silently alter its input.

### Select the accepted basis durably

The first day has nothing to carry forward, so it searches. Every later day
starts from the record ids the previous run accepted and adds whatever Graft
reports as accepted since the previous snapshot. The basis therefore grows
only through explicit acceptance; a keyword search that happens to match a
different set of records on day 40 cannot change it.

```{r basis}
basis_path <- "briefing/basis.rds"
accepted_classes <- c("Claim", "ClaimSupport", "EvidenceSpan", "Source")

if (file.exists(basis_path)) {
  previous <- readRDS(basis_path)
  changed <- graft_changes(
    view,
    since = previous$snapshot,
    class = accepted_classes
  )
  record_ids <- union(previous$record_ids, changed$record_id)
} else {
  claims <- graft_find(
    view,
    "grid-scale battery recycling",
    class = "Claim",
    limit = 25
  )
  record_ids <- claims$id
}

knowledge <- tempest_knowledge(view, record_ids = record_ids)
```

`graft_changes()` compares two committed boundaries store-wide and returns one
row per changed record with its action, the union of changed public fields,
and the latest accepted record. A `GraftView` bounds it at the pinned
snapshot, so the selection cannot include a commit that landed after the run
started.

## Research today's change

Use a small expert panel that reflects the decision, not a collection of
generic personas. The topic asks for the briefing's decision structure while
Tempest keeps the exact report, sources, claims, and claim-support pairs behind
its ordinary product interface.

```{r research}
technical <- tempest_expert(
  name = "Technical readiness reviewer",
  title = "Battery recycling process specialist",
  description = "Reviews demonstrated process performance and scale-up risk.",
  instructions = paste(
    "Separate measured performance from projections.",
    "Call out consequential unknowns. Report no change only when a verified",
    "claim restates a condition the organization already accepted."
  ),
  focus_areas = c("process yield", "scale-up", "safety")
)

market <- tempest_expert(
  name = "Market and policy reviewer",
  title = "Battery supply-chain analyst",
  description = "Reviews policy, capacity, partnerships, and market movement.",
  instructions = paste(
    "Distinguish announcements from operating evidence.",
    "State which developments could change a decision today."
  ),
  focus_areas = c("policy", "capacity", "partnerships", "economics")
)

result <- tempest_run(
  topic,
  config = tempest_config(
    models = "openai/gpt-5.6-luna",
    citation_policy = "claim_verified",
    on_unsupported_claim = "flag"
  ),
  knowledge = knowledge,
  experts = list(technical, market),
  max_questions_per_perspective = 2,
  output_dir = "briefing/runs",
  run_id = paste0("daily-", Sys.Date())
)

report <- tempest_report(result)
sources <- tempest_sources(result)
claims <- tempest_claims(result)
supports <- tempest_claim_supports(result)
```

Treat the report as the briefing and the tables as its evidence rail. The
at-a-glance brief and each detailed section keep four statements visibly
distinct, and the first and last of them are decided against the pinned
snapshot rather than by the prompt:

- "What changed" items copy verified claims exactly, and Tempest accepts one
  only when its text does not restate a `Claim` already in the pinned
  knowledge. On the first day every verified claim is new.
- "Why it matters" assessments are labeled, confidence-calibrated inferences.
- "Review today" items are proposed actions.
- "No material change" statements copy one verified claim exactly, and
  Tempest accepts one only when that claim restates an accepted `Claim`. Today's
  evidence re-verified something the organization already knew.

Missing, unresolved, or inconclusive evidence does not become a no-change
statement, because it does not produce a verified claim at all. Tempest binds
every item to its exact claim IDs in machine-readable provenance, re-checks
each item's disposition against the accepted claims whenever the report is
rendered, and rejects a report if that provenance no longer matches the
visible Markdown.

Whether somebody was shown the briefing or read it is operational state; it is
not accepted knowledge. An assessment or review action is also not accepted
knowledge merely because it appears in the briefing.

## Review the evidence and the run separately

Prepare a Graft plan, but do not commit it. In parallel, project the completed
run into scans. The first review asks whether the proposed knowledge is fit to
accept; the second asks whether the agent trajectory itself behaved well.

```{r review}
proposal <- tempest_promotion_bundle(result)
plan <- tempest_graft_plan(store, proposal)

review <- tempest_trajectory_review(
  result,
  promotion_bundle = proposal
)
trajectory <- scans::as_trajectory_tempest(review)

run_summary <- scans::summarize_trajectories(trajectory)
run_findings <- scans::scan_trajectories(trajectory)

plan@valid
plan@changes[, c("class", "record_id", "action", "disposition", "changed_fields")]
run_summary
run_findings
```

Tempest keys accepted `Claim` identity on the claim's normalized statement
text and accepted `Source` identity on its locator. A claim the organization
already accepted therefore plans as an `update` whose `changed_fields` name
only run metadata such as the Tempest claim id and verification time, and
Graft's `disposition` column reads `revision`; a genuinely new statement reads
`new`; an identical proposal reads `duplicate`. A staged claim that sets
`superseded_by`, or evidence whose `support_type` is `contradicts`, surfaces
as `superseded`, `supersedes`, or `contradicted` so the reviewer sees
statement-level relations without reading every row.

Do not collapse these into one confidence score. Claim support, proposed Graft
changes, and trajectory findings answer different review questions.

## Accept deliberately, including no change

Accepted knowledge changes only when a `Claim` is new, superseded, or
contradicted. A plan whose claims are all `duplicate` or `revision` proposes
nothing the organization does not already know; publish the briefing and
record that no-change outcome in the host's run log rather than manufacturing a
knowledge revision to prove the schedule ran.

When the reviewer accepts a real change, commit the exact plan that was shown,
bind a Tempest receipt to the resulting immutable snapshot, and save the
basis the next run will carry forward:

```{r accept}
claim_rows <- plan@changes[plan@changes$class == "Claim", ]
has_change <- any(
  claim_rows$disposition %in% c("new", "supersedes", "superseded", "contradicted")
)

if (has_change) {
  commit_result <- graft_commit(store, plan)
  receipt <- tempest_promotion_receipt(
    store,
    proposal,
    plan,
    commit_result
  )

  accepted_review <- tempest_trajectory_review(
    result,
    promotion_bundle = proposal,
    promotion_receipt = receipt
  )

  accepted_ids <- vapply(
    Filter(
      \(revision) revision$class %in% accepted_classes,
      receipt@record_revisions
    ),
    `[[`,
    character(1),
    "record_id"
  )
  saveRDS(
    list(
      snapshot = graft_snapshot(store),
      record_ids = union(record_ids, accepted_ids)
    ),
    basis_path
  )
}
```

Rejection leaves accepted knowledge unchanged, and the saved basis keeps
pointing at the previous snapshot. This two-speed loop keeps routine monitoring
passive and makes only material changes enter the slower decision-and-review
path.

## Let a person edit accepted knowledge

The store was opened with `okf = "managed"`, so Graft keeps a readable Open
Knowledge Format working tree next to the DuckDB file. A reviewer who wants to
correct a wording, retract a claim, or mark one statement as superseding
another edits those files; Graft turns the edits back into the same kind of
plan the briefing produced, and the same acceptance step applies.

```{r okf}
graft_sync(store)
graft_status(store)$status
#> "current"

# A person edits files under the managed OKF directory, then:
graft_status(store)$status
#> "modified"

edited <- graft_review(
  store,
  provenance = graft_provenance(
    "briefing-reviewer",
    idempotency_key = paste0("okf-", Sys.Date())
  )
)
edited@changes[, c("class", "record_id", "action", "disposition")]

if (edited@valid) {
  graft_commit(store, edited)
  graft_sync(store)
}
```

The edited plan is bound to the exact bundle and accepted batch it was
reviewed against, so a commit that landed in between fails the plan instead
of silently rebasing the person's edits. Tomorrow's briefing picks the edit up
through `graft_changes()` like any other accepted change.

```{r close}
graft_close(store)
```
