---
title: "Research, review, retain, and reuse"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Research, review, retain, and reuse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Tempest brings the Ellmerverse Plus packages together in a research workflow:
deputy runs bounded agents, dsprrr executes typed scientific programs, shinymcp
presents an exact result for inspection, and graft preserves reviewed evidence
for another session. Tempest owns research state and scientific validation.

Tempest requires deputy 0.0.0.9002 or later. Research tools declare their actual
effects: local evidence inspection is read-only and closed-world, adding a
proposed claim changes only provisional research, and web retrieval also writes
retrieved evidence to that workspace and is open-world. Neither operation is
destructive. Moderator delegation can change provisional research and reach
the web. The adapter allowlists its tools while keeping file access, shell and R
execution, and package installation disabled. It does not override Deputy's base
permission checks. Plan mode refuses provisional claim writes, web retrieval,
and delegation even when allowlisted. Readonly mode requires custom tools to
be allowlisted; a host offering evidence inspection should allowlist only the
read tools.

## Try the complete workflow

Install the demonstration's optional packages alongside Tempest:

```{r, eval=FALSE}
pak::pak(c("JamesHWade/tempest", "JamesHWade/shinymcp", "JamesHWade/graft",
           "shiny", "bslib", "callr"))
tempest::tempest_app(demo = TRUE)
```

No model credentials or network requests are needed after installation. Scripted
Chat responses and local synthetic evidence run through the actual deputy,
dsprrr, and Tempest implementations. The demonstration proves composition and
evidence integrity; it does not measure live-model accuracy, cost, or latency.
The replay requests registered evidence tools through Deputy's tool callbacks,
so permissions are checked before those functions execute.
Research, retention, and reuse run in background R processes so the interface
stays responsive. Before acceptance, the worker restores the saved run and
checks that its exact review identity still matches the candidate on screen.

1. Inspect the initial report in the **Research card**. Open **Evidence**, choose
   the claim, and read its exact source quote, support judgment, and source locator.
2. Check that you reviewed this candidate and give a reason. **Accept reviewed
   research** records a Graft decision for the complete report and evidence.
3. **Reopen in a fresh R process** rechecks current acceptance, admits the evidence
   into a new Co-STORM session, and verifies exact source equality after save/resume.
4. **Research the corrected note** consults the current accepted evidence and
   produces a separate 74% recovery candidate. The accepted 82% result remains
   current until you explicitly review and accept the correction.
5. **Withdraw retained research**, then attempt reuse again. Reuse is refused;
   historical reports and the decision history remain readable.

Every number and scientific statement in the example is invented. The source
locators use `example.org`; no source is fetched from that domain.

The default demo files live under R's temporary directory. To keep the evidence
and completed runs across R sessions, choose a local directory before launching:

```{r, eval=FALSE}
options(tempest.demo_dir = "~/tempest-ecosystem-demo")
tempest::tempest_app(demo = TRUE)
```

The demo is a trusted local, single-user application. Its eligibility callback
admits only accepted decisions from `local-demo-reviewer` in the `synthetic-pilot`
stream for `demo-briefing`. This is a fixed example policy. Applications with
multiple users must bind their own authenticated reviewer and current access policy.

## Put your own research in a card

```{r, eval=FALSE}
result <- tempest::tempest_run("Battery recovery at pilot scale")
card <- tempest::tempest_mcp_app(result)
shinymcp::preview_app(card)

# Serve in a compatible MCP client, or use mcp_endpoint() for deployment.
shinymcp::serve(card)
```

The same constructor accepts a succeeded, quiescent Co-STORM session after
`session$publish()`. It reads the existing report, evidence, and trajectory once.
Subsequent research in that session does not rewrite the card; build a new card
to review a newer completed product. Rendering or restoring a card never reruns
research or records acceptance.
Each exact review gets its own MCP resource name, keeping a host's cached page
consistent with the new report and claim tools when a candidate changes.

Clients without interactive pages can use the same three read-only tools:

| Tool | Result |
|---|---|
| `review_tempest_research` | Compact summary, exact review identity, and claim index |
| `inspect_tempest_claim` | One claim and its exact source quotes, locations, and support assessments |
| `read_tempest_report` | The exact committed Markdown report |

The full rendered report and presentation tables travel as page metadata. The
model gets a compact overview and can request the evidence it needs. Report
rendering keeps safe Markdown elements and attributes, removes active HTML and
embedded images, and suppresses links other
than HTTP, HTTPS, and in-page anchors. The exact stored Markdown is unchanged.

In a Shiny host, create the card inside a render function:

```{r, eval=FALSE}
output$research_card <- shiny::renderUI({
  shinymcp::mcp_embed(
    tempest::tempest_mcp_app(completed_research()),
    tool = "review_tempest_research"
  )
})
```

## Keep decisions in the host

The card is read-only. Acceptance, correction, and withdrawal use the existing
Graft operations from a host control that binds the exact reviewed product,
reviewer, purpose, and expected decision. A completed workflow or a positive
verifier score does not count as a human review. Model tool annotations do not
authorize an application effect.

```{r, eval=FALSE}
store <- graft::graft_store("research-evidence", create = TRUE)
selection <- tempest::tempest_publish_artifact_research(result, store)
# Called only after the host reviewer checks this exact candidate.
decision <- graft::graft_accept(
  store, graft::graft_read_selection(store, selection), "pilot",
  expected = NULL, key = "review-1", actor = "authenticated-reviewer",
  reason = "Reviewed source evidence", purpose = "research-briefing"
)
knowledge <- tempest::tempest_reuse_artifact_research(
  store, "pilot", decision@id, "research-briefing",
  eligible = function(event) host_can_consult(event)
)
```

Use the decision observed when the review began as `expected`. Refresh and
review again if a newer decision makes that basis stale. The demo resets its
review checkbox when the candidate or review basis changes. A correction is a
new proposal with an exact input basis; acceptance is never automatic.

## Verify the composition

From a package checkout:

```r
testthat::test_local(
  filter = "^(tools|deputy-adapter|costorm-deputy|ecosystem-contracts|mcp-app|ecosystem-demo)$"
)
```

These deterministic checks exercise Deputy permissions, actual moderator and
expert delegation, Deputy execution, and dsprrr stage
identities, both model and page reads, exact claim evidence, stale acceptance,
proposal/acceptance separation, correction with accepted input, a fresh-process
restart, historical reads, and refusal of cached acceptance after withdrawal.
The supported CI lane installs the dependency revisions pinned in `DESCRIPTION`;
a separate upstream lane checks current default branches without changing the
supported set.
