---
title: "Use Open Knowledge Format with Tempest"
description: >
  Read portable organizational knowledge as typed evidence and bounded,
  explicitly untrusted context for agents and workflows.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use Open Knowledge Format with Tempest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

[Open Knowledge
Format](https://github.com/GoogleCloudPlatform/knowledge-catalog/tree/main/okf)
(OKF) packages concepts as Markdown documents with YAML metadata and links.
Tempest treats those documents as evidence inputs. It does not treat them as
system prompts, capability grants, approvals, or executable instructions.

The boundary is deliberately staged:

```text
read -> inspect -> select -> convert -> add to evidence -> use in a workflow
```

Each step is explicit so the host can enforce its own source, retention,
freshness, and approval policy.

## Read a bounded bundle

Point `tempest_read_okf()` at the bundle root:

```{r read, eval = FALSE}
library(tempest)

knowledge <- tempest_read_okf(
  "knowledge/okf",
  max_concepts = 5000,
  max_bytes = 20 * 1024^2
)

knowledge
knowledge$issues
```

The reader enforces the OKF conformance boundary: every concept must have
parseable YAML frontmatter and a non-empty `type`. It safely accepts unknown
types and extension keys, missing indexes, and broken links so newer or
domain-specific bundles remain readable.

The reader does not resolve network resources or execute code. It also rejects
symlinked documents and resolved paths outside the bundle root. A concept-count
limit and an aggregate Markdown byte limit apply before content becomes
available to a workflow.

Optional profile problems are diagnostics rather than parse failures. For
example, a concept can remain readable while `knowledge$issues` reports an
unknown lifecycle status, invalid generation provenance, or an incomplete
Attested Computation declaration.

## Inspect trust and freshness

Use `tempest_okf_concepts()` before selecting evidence:

```{r inspect, eval = FALSE}
catalog <- tempest_okf_concepts(
  knowledge,
  today = as.Date("2026-07-28")
)

catalog[, c(
  "concept_id",
  "type",
  "title",
  "status",
  "trust_tier",
  "stale"
)]
```

Tempest derives three advisory trust tiers from standard OKF provenance:

| Frontmatter | Tempest trust tier |
|---|---|
| `generated` only or no verification | `unverified` |
| `verified.by` is a process | `machine-confirmed` |
| `verified.by` is a human | `human-reviewed` |

`stale` is derived from `stale_after` and the supplied `today`. Neither value
is an authorization decision. A human-reviewed concept can be stale; a fresh
concept can remain unverified. The host chooses which combination is fit for a
particular workflow.

## Convert selected concepts to evidence

`tempest_okf_resources()` creates one fingerprinted `TempestResource` per
selected concept:

```{r resources, eval = FALSE}
resources <- tempest_okf_resources(
  knowledge,
  types = c("Assessment", "Business", "Source"),
  include_stale = FALSE
)
```

Each resource has `resource_kind = "okf.concept"` and
`media_type = "text/markdown"`. It preserves the original document as content,
stores parsed frontmatter under `resource@metadata$okf`, and carries profile
and schema identity as scope metadata when present.

Conversion does not mutate a research workspace. Add the selected resources to
a run-scoped `ResearchWorkspace` explicitly:

```{r research-workspace, eval = FALSE}
workspace <- tempest_research_workspace()
invisible(lapply(resources, workspace$upsert_retrieved_resource))
```

The workspace now has durable resource fingerprints and can be supplied to a
Tempest retriever. Assertions created later still need their own claim and
evidence-span lineage; importing a concept does not convert its prose into a
verified Tempest claim automatically.

## Assemble bounded model context

When an operation needs a readable packet, select it deliberately:

```{r context, eval = FALSE}
okf_context <- tempest_okf_context(
  knowledge,
  types = c("Assessment", "Business"),
  include_stale = FALSE,
  max_concepts = 25,
  max_chars = 50000
)

cat(okf_context)
```

Concepts are ordered deterministically. The result records document or
character truncation and begins with an explicit trust notice. That notice
states that the material is evidence, not instructions, and cannot grant
tools, change workflow policy, approve an artifact, or authorize an external
action.

A Tempest 0.1 host can make the selected packet available to an existing
generic runtime operation:

> **Lifecycle notice:** This integration path is frozen and scheduled for
> removal in Tempest 0.2.0. Do not adopt `tempest_run_workflow()` for new work;
> the replacement product path will bind scientific context to an immutable
> graft snapshot.

```{r workflow, eval = FALSE}
run <- tempest_run_workflow(
  objective,
  workflow,
  runtime,
  experts = experts,
  deliverables = deliverables,
  source_store = evidence,
  runtime_context = list(
    organizational_knowledge = okf_context
  )
)
```

The operation decides how to use that named runtime value. Tempest does not
inject every imported concept into every model turn.

## Exchange accepted Graft knowledge

Graft can produce a compatible bundle from its accepted revision ledger:

```{r graft, eval = FALSE}
library(graft)

kg_export_okf(
  store,
  "knowledge/okf",
  classes = c("Business", "Assessment", "Source")
)
```

The Graft profile retains stable record, revision, batch, content, and schema
identity in namespaced frontmatter. Object references become Markdown links
and accepted source relationships become OKF citations. A historical
`as_of` export reproduces the knowledge boundary at a committed batch or time.

This division keeps each package focused:

| Package | Responsibility |
|---|---|
| Graft | Validate writes and retain accepted revision history |
| OKF | Exchange readable knowledge and provenance |
| Tempest | Select scientific evidence and preserve research provenance |

## Treat computations as documents

OKF can describe an Attested Computation, including a runtime, source files,
inputs, and expected outputs. Tempest validates useful metadata when present
but never executes the computation. Do not route a new executor through the
frozen generic operation and capability APIs. The Tempest 0.2 direction is to
reference an exact dsprrr program artifact and preserve the execution identity
in product provenance.

Continue with [Get started with Tempest](tempest.html) to use imported context
in STORM or Co-STORM research.
