---
title: "Get started with Tempest"
description: >
  Choose a Tempest workflow, configure a provider, create an evidence-backed
  report, and continue into interactive scientific research.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with Tempest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Tempest brings STORM and Co-STORM research workflows to R. It combines
multi-perspective research, source and claim tracking, cited writing, selected
or generated experts, scientific reports, and resumable research state.

This article helps you choose a workflow and produce a first result.
Provider-backed examples are not executed while the vignette is built, so
rendering the article makes no model or search calls and requires no provider
credentials.

## Choose a workflow

Start with the interface that matches the outcome you need:

| Goal | Interface | Result |
|---|---|---|
| Create an evidence-backed report | `tempest_run()` | A report, sources, claims, outline, and run state |
| Explore a topic interactively | `run_app()` or `tempest_session()` | A continuing Co-STORM conversation, mind map, evidence, and report |
| Inspect the frozen generic kernel | `tempest_run_workflow()` | Frozen experimental 0.1 behavior scheduled for removal in 0.2.0 |

Most package users should begin with scripted STORM. Co-STORM is useful when
the research question should evolve through dialogue. The experimental generic
workflow kernel is frozen; do not use it for new integrations.

## Install and load Tempest

Install the development package from GitHub:

```{r installation, eval = FALSE}
# install.packages("pak")
pak::pak("JamesHWade/tempest")
```

Then load it:

```{r library}
library(tempest)
```

## Configure models and search

Tempest uses `ellmer` for model access. Configure the credential required by
your chosen provider outside your R source files. For the default OpenAI
models, Tempest uses `ellmer::chat_openai(auth = "codex")` and reuses the
file-backed ChatGPT subscription authentication managed by Codex CLI. If Codex
has not stored file-backed credentials, run
`codex login -c 'cli_auth_credentials_store="file"'`. Do not put credentials in
a Tempest configuration, expert profile, workflow specification, or saved run.

`tempest_config()` controls model roles, retrieval, evidence policy, caching,
and Co-STORM behavior. A single model string can be used for every role, or
`models` can be a named list for finer control.

This small first-run configuration uses Wikipedia search, which does not need
a separate search-provider key:

```{r config}
cfg <- tempest_config(
  models = "openai/gpt-5.6-luna",
  search_provider = "wikipedia",
  max_search_results = 4,
  max_sources = 12,
  citation_policy = "source_attributed"
)
```

The default `"native"` search provider uses provider-native web search when
available and falls back to Wikipedia. Other providers may require their own
environment variables.

## Create a first STORM report

`tempest_run()` performs five stages:

1. discover perspectives and expert profiles;
2. research questions from each perspective;
3. draft and refine an outline;
4. write cited sections; and
5. polish and publish the Markdown report.

Start with two experts and a small question budget while learning the package:

```{r first-run, eval = FALSE}
result <- tempest_run(
  "What are the main technical and policy barriers to recycling grid-scale batteries?",
  config = cfg,
  n_experts = 2,
  max_questions_per_perspective = 2,
  output_dir = "tempest-runs",
  run_id = "grid-battery-recycling",
  verbose = TRUE
)
```

This call uses the configured model and search provider, so it incurs provider
usage and requires network access. `output_dir` creates a run directory and
persists each stage after it completes.

The returned list contains the main report and the state used to create it:

```{r inspect-result, eval = FALSE}
cat(result$report_md)

result$title
result$perspectives
result$experts
result$outline

sources <- tempest_sources(result$store)
claims <- tempest_claims(result$store)

utils::head(sources[c("id", "title", "url")])
utils::head(claims[
  c(
    "claim_id",
    "claim_text",
    "source_ids",
    "verification_status",
    "support_score"
  )
])
```

The source store holds run-scoped provisional evidence. Sources, claims,
support scores, and citations remain inspectable instead of being flattened
into report text. The product report is available directly in
`result$report_md`; new callers should not depend on the frozen 0.1 artifact
catalog.

Choose a stronger `citation_policy` when a workflow requires verified claims:

- `"source_attributed"` converts known inline source IDs to footnotes without
  verifying the claims;
- `"claim_verified"` runs claim verification and adds verification status to
  the final references; and
- `"strict"` also applies the configured unsupported-claim behavior to inline
  citations.

Use `min_support_score` in `tempest_config()` to set the verification threshold.
Under the `"strict"` policy, `on_unsupported_claim` controls whether weak
claims are flagged, dropped, revised, or retained with a warning.

## Resume a staged run

When `output_dir` is supplied, Tempest saves completed stages, source and claim
records, report artifacts, and checksummed metadata beneath the run directory.
Resume with the same topic, configuration, output directory, and run ID:

```{r resume, eval = FALSE}
result <- tempest_run(
  "What are the main technical and policy barriers to recycling grid-scale batteries?",
  config = cfg,
  n_experts = 2,
  max_questions_per_perspective = 2,
  output_dir = "tempest-runs",
  run_id = "grid-battery-recycling",
  resume = TRUE,
  verbose = TRUE
)
```

Completed stages are loaded rather than rerun. Keep model, retrieval, and
workflow settings stable when continuing an existing run.

## Supply your own expert team

By default, STORM generates a topic-specific expert pool and pairs it with the
discovered perspectives. A host can instead choose exact profiles from its own
expert pool:

```{r selected-experts, eval = FALSE}
experts <- list(
  tempest_expert(
    expert_id = "expert.recycling-engineering",
    name = "Recycling Engineer",
    title = "Battery recovery specialist",
    description = "Focuses on process yield, safety, and scale-up.",
    instructions = "Separate demonstrated performance from projections."
  ),
  tempest_expert(
    expert_id = "expert.recycling-policy",
    name = "Policy Analyst",
    title = "Circular-economy policy specialist",
    description = "Focuses on incentives, standards, and accountability.",
    instructions = "Compare jurisdictions and preserve policy uncertainty."
  )
)

result <- tempest_run(
  "What are the main technical and policy barriers to recycling grid-scale batteries?",
  config = cfg,
  experts = experts,
  output_dir = "tempest-runs",
  run_id = "selected-expert-recycling",
  verbose = TRUE
)
```

Profiles describe a scientific perspective and stable expert identity. Keep
live tools, clients, and credentials outside the serialized profile. The 0.1
skill, capability, and connection fields are frozen and should not be used for
new integrations.

## Explore interactively with Co-STORM

Launch the bundled Shiny application for a multi-expert conversation, live
mind map, evidence tables, and report generation. The app uses optional
packages from `Suggests`; install them if they are not already available:

```{r app, eval = FALSE}
pak::pak(
  c(
    "shiny",
    "bslib",
    "shinychat",
    "DT",
    "visNetwork",
    "promises",
    "later",
    "mirai",
    "zip"
  )
)

run_app()
```

For console control, create a session directly:

```{r costorm, eval = FALSE}
session <- tempest_session(
  "Grid-scale battery recycling",
  config = cfg,
  n_experts = 3
)

session$warmup(verbose = TRUE)
answer <- session$step(
  "Which barriers are most likely to constrain deployment this decade?"
)
cat(answer$answer)

report <- session$report(
  style = "executive",
  include_references = TRUE
)
cat(report)

tempest_session_save(
  session,
  "tempest-runs/grid-battery-session",
  overwrite = TRUE
)

restored <- tempest_session_resume(
  "tempest-runs/grid-battery-session",
  config = cfg
)
```

Each expert keeps its own conversation continuity and receives only the
role-specific tools needed for that session. The moderator delegates by stable
expert ID, and evidence collected during dialogue remains available to the
final report. Session bundles preserve durable research state, but not
credentials, live chat handles, tools, or authenticated clients. Recreate those
host-injected dependencies after resume. Existing Tempest 0.1 integrations that
supplied a custom `runtime` must still reattach it; that seam is frozen and
should not be used for new integrations.

## Maintain a Tempest 0.1 application-specific workflow

> **Lifecycle notice:** The experimental generic workflow kernel is frozen and
> scheduled for removal in Tempest 0.2.0. Do not use it for new integrations.

Existing Tempest 0.1 hosts may still need the generic kernel while the product
migration is under way.

The `reusable-workflows` vignette documents maintenance of an existing Tempest
0.1 integration with:

- an application-neutral objective;
- a host-selected expert;
- a versioned output template;
- runtime generator, validator, renderer, and step operations;
- artifact approval and ordered events;
- typed output inspection; and
- save/restore with explicit runtime reattachment.

Open it from an installed package:

```{r reusable-vignette, eval = FALSE}
vignette("reusable-workflows", package = "tempest")
```

On the pkgdown website, see
[Build reusable workflows with Tempest](reusable-workflows.html).

## Where to go next

- Use `tempest_run()` for a scripted STORM report.
- Use `tempest_session()` or `run_app()` for interactive Co-STORM research.
- Add semantic retrieval with `ragnar` through `embed_fn` or a pre-built
  `ragnar_store`.
- Optimize structured STORM modules with
  `tempest_optimize_dsprrr_modules()`.
- Capture progress with `tempest_progress_collector()` and inspect events with
  `tempest_progress_event_data()`.
- Evaluate scripted and interactive workflows with `tempest_task()` and
  `tempest_costorm_task()`.

Start with a narrow topic, a small expert and question budget, and durable
output. Expand model roles, retrieval, parallelism, and evaluation only after
the first workflow behaves as intended.
