---
title: "Get started with Tempest"
description: >
  Choose a Tempest workflow, configure a provider, create an evidence-backed
  report, and continue into interactive scientific research.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with Tempest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Tempest brings STORM and Co-STORM research workflows to R. It combines
multi-perspective research, source and claim tracking, cited writing, selected
or generated experts, scientific reports, and resumable research state.

This article helps you choose a workflow and produce a first result.
Provider-backed examples are not executed while the vignette is built, so
rendering the article makes no model or search calls and requires no provider
credentials.

## Choose a workflow

Start with the interface that matches the outcome you need:

| Goal | Interface | Result |
|---|---|---|
| Create an evidence-backed report | `tempest_run()` | A report, sources, claims, outline, and run state |
| Explore a topic interactively | `run_app()` or `tempest_session()` | A continuing Co-STORM conversation, mind map, evidence, and report |

Most package users should begin with scripted STORM. Co-STORM is useful when
the research question should evolve through dialogue. Tempest 0.2 has no
application-neutral workflow or runtime API.

## Install and load Tempest

Install the development package from GitHub:

```{r installation, eval = FALSE}
# install.packages("pak")
pak::pak("JamesHWade/tempest")
```

Then load it:

```{r library}
library(tempest)
```

## Configure models and search

Tempest uses `ellmer` for model access. Configure the credential required by
your chosen provider outside your R source files. For the default OpenAI
models, Tempest uses `ellmer::chat_openai(auth = "codex")` and reuses the
file-backed ChatGPT subscription authentication managed by Codex CLI. If Codex
has not stored file-backed credentials, run
`codex login -c 'cli_auth_credentials_store="file"'`. Do not put credentials in
a Tempest configuration, expert profile, workflow specification, or saved run.

`tempest_config()` controls model roles, retrieval, evidence policy, caching,
and Co-STORM behavior. A single model string can be used for every role, or
`models` can be a named list for finer control.

This small first-run configuration uses Wikipedia search, which does not need
a separate search-provider key:

```{r config}
cfg <- tempest_config(
  models = "openai/gpt-5.6-luna",
  search_provider = "wikipedia",
  max_search_results = 4,
  max_sources = 12,
  citation_policy = "source_attributed"
)
```

The default `"native"` search provider uses provider-native web search when
available and falls back to Wikipedia. Other providers may require their own
environment variables.

## Create a first STORM report

`tempest_run()` performs five stages:

1. discover perspectives and expert profiles;
2. research questions from each perspective;
3. draft and refine an outline;
4. write cited sections; and
5. polish and publish the Markdown report.

Start with two experts and a small question budget while learning the package:

```{r first-run, eval = FALSE}
result <- tempest_run(
  "What are the main technical and policy barriers to recycling grid-scale batteries?",
  config = cfg,
  n_experts = 2,
  max_questions_per_perspective = 2,
  output_dir = "tempest-runs",
  run_id = "grid-battery-recycling",
  verbose = TRUE
)
```

This call uses the configured model and search provider, so it incurs provider
usage and requires network access. `output_dir` creates a run directory and
persists each stage after it completes.

The returned list contains the main report and the state used to create it:

```{r inspect-result, eval = FALSE}
cat(result$report_md)

result$title
result$perspectives
result$experts
result$outline

sources <- tempest_sources(result$workspace)
claims <- tempest_claims(result$workspace)
supports <- tempest_claim_supports(result$workspace)

utils::head(sources[c("id", "title", "url")])
utils::head(claims[
  c(
    "claim_id",
    "claim_text",
    "source_ids",
    "verification_status",
    "support_score"
  )
])
utils::head(supports)
```

The research workspace holds run-scoped provisional evidence. Retrieved
resources, proposed claims, exact claim-by-evidence-span support records, and
citations remain inspectable instead of being flattened into report text. The
pair records are authoritative; claim summaries and citation-audit tables are
derived projections. The product report is available directly in
`result$report_md`; new callers should not depend on the frozen 0.1 artifact
catalog.

Every STORM and Co-STORM publication runs the exact verifier ProgramSet stage
and atomically binds claim-by-evidence-span support. `citation_policy` controls
only report rendering and unsupported-claim handling:

- `"source_attributed"` converts known inline source IDs to footnotes without
  disabling product verification;
- `"claim_verified"` exposes threshold-verified support in the rendered report
  and final references; and
- `"strict"` fails publication unless each publishable assertion is bound to a
  completed, provenance-bound verification result at the configured threshold
  and cites that claim's exact source set.

Use `min_support_score` in `tempest_config()` to set the verification threshold.
Under the `"strict"` policy, `on_unsupported_claim` controls whether weak
claims are flagged, dropped, revised, or retained with an explicit warning.
They are never presented as unmarked verified facts.

## Review promotion to Graft

A succeeded product can become a deterministic proposal for Graft review
without granting Tempest acceptance authority:

```{r promotion, eval = FALSE}
bundle <- tempest_promotion_bundle(
  workspace = result$workspace,
  manifest = result$manifest,
  stage_records = result$state$stage_records
)
trusted_bundle_id <- bundle@bundle_id
tempest_save_promotion_bundle(bundle, "promotion/grid-battery-recycling")
bundle <- tempest_read_promotion_bundle(
  "promotion/grid-battery-recycling",
  expected_bundle_id = trusted_bundle_id
)

schema <- tempest_graft_schema()
# Open `store` with this schema using the host's chosen Graft location.
plan <- tempest_graft_plan(store, bundle)

# Review the plan before the host explicitly accepts it.
commit_result <- graft::graft_commit(store, plan)
receipt <- tempest_promotion_receipt(store, bundle, plan, commit_result)
```

Tempest does not commit during bundle construction or planning. The current
closed promotion format contains exact Sources, Claims, EvidenceSpans,
ClaimSupports, and extraction and verification ProgramArtifacts. A closed
proof projection retains the exact resources, claims, spans, and supports
needed to recompute every retained StageRecord digest. A selection must include
every output bound by each retained extraction or verification record; Tempest
rejects partial stage-output selection instead of packaging nonselected
evidence. Reading requires the original bundle id as an out-of-band trust pin;
bundle-local checksums establish internal consistency, not authenticity. Older
bundle shapes are rejected. Research promotion never mints a
`GovernedProcedure`; that record requires its own reviewed Graft acceptance
flow.

## Resume a staged run

When `output_dir` is supplied, Tempest saves completed stages, source and claim
records, report artifacts, and checksummed metadata beneath the run directory.
Resume with the same topic, configuration, output directory, and run ID:

```{r resume, eval = FALSE}
result <- tempest_run(
  "What are the main technical and policy barriers to recycling grid-scale batteries?",
  config = cfg,
  n_experts = 2,
  max_questions_per_perspective = 2,
  output_dir = "tempest-runs",
  run_id = "grid-battery-recycling",
  resume = TRUE,
  verbose = TRUE
)
```

Completed stages are loaded rather than rerun. Keep model, retrieval, and
workflow settings stable when continuing an existing run.

Current readers accept only `ResearchWorkspace` snapshot schema 5, Co-STORM
snapshot and bundle schema 9, STORM bundle schema 7 with state schema 4,
ProgramSet schema 2, research-manifest schema 3, StageRecord output-digest payload
schema 3, and promotion-bundle schema 1. Every other version is rejected, as is
any missing or extra field or value that becomes valid only after coercion.

Every typed attempt is saved with its exact stage, program, evaluator, trace,
support decision, and fallback path. Structured output is validated as a whole
before product state changes. A resumed running attempt is recorded as
cancelled, and the final report adds a deterministic `Execution review` for
failed, cancelled, policy-fallback, or nonverified grounded outcomes.

## Supply your own expert team

By default, STORM generates a topic-specific expert pool and pairs it with the
discovered perspectives. A host can instead choose exact profiles from its own
expert pool:

```{r selected-experts, eval = FALSE}
experts <- list(
  tempest_expert(
    expert_id = "expert.recycling-engineering",
    name = "Recycling Engineer",
    title = "Battery recovery specialist",
    description = "Focuses on process yield, safety, and scale-up.",
    instructions = "Separate demonstrated performance from projections."
  ),
  tempest_expert(
    expert_id = "expert.recycling-policy",
    name = "Policy Analyst",
    title = "Circular-economy policy specialist",
    description = "Focuses on incentives, standards, and accountability.",
    instructions = "Compare jurisdictions and preserve policy uncertainty."
  )
)

result <- tempest_run(
  "What are the main technical and policy barriers to recycling grid-scale batteries?",
  config = cfg,
  experts = experts,
  output_dir = "tempest-runs",
  run_id = "selected-expert-recycling",
  verbose = TRUE
)
```

Profiles describe a scientific perspective and stable expert identity. Keep
live tools, clients, and credentials outside the serialized profile. Product
code attaches only the tools required for its STORM or Co-STORM role.

## Explore interactively with Co-STORM

Launch the bundled Shiny application for a multi-expert conversation, live
mind map, evidence tables, and report generation. The app uses optional
packages from `Suggests`; install them if they are not already available:

```{r app, eval = FALSE}
pak::pak(
  c(
    "shiny",
    "bslib",
    "shinychat",
    "DT",
    "visNetwork",
    "promises",
    "later",
    "mirai",
    "zip"
  )
)

run_app()
```

For console control, create a session directly:

```{r costorm, eval = FALSE}
session <- tempest_session(
  "Grid-scale battery recycling",
  config = cfg,
  n_experts = 3
)

session$warmup(verbose = TRUE)
answer <- session$step(
  "Which barriers are most likely to constrain deployment this decade?"
)
cat(answer$answer)

report <- session$report(
  style = "executive",
  include_references = TRUE
)
cat(report)

tempest_session_save(
  session,
  "tempest-runs/grid-battery-session",
  overwrite = TRUE
)

restored <- tempest_session_resume(
  "tempest-runs/grid-battery-session",
  config = cfg
)
```

Each expert keeps its own conversation continuity and receives only the
role-specific tools needed for that session. The moderator delegates by stable
expert ID, and evidence collected during dialogue remains available to the
final report. Session bundles preserve durable research state, but not
credentials, live chat handles, tools, or authenticated clients. Configure
fresh supported chats and retrieval dependencies through `config` before
resuming.

The moderator and experts use persistent Deputy agents as the required
Co-STORM runtime. Tempest disables ambient file, shell, R, web, and
package-install capabilities, then allowlists only the tools already attached
for that role. Snapshots persist canonical opaque terminal traces for each run
and never serialize the Deputy Agent or provider credentials. Those execution
identities support correlation and audit joins only; they do not claim that an
execution caused, authored, or validated report content.

## Generic-kernel cutover

Tempest 0.2 supports only `tempest_run()` and `tempest_session()` as research
product entry points. The experimental application-neutral kernel is
unavailable, and retained symbols fail immediately with
`tempest_generic_kernel_cutover_error`. No compatibility layer is provided.

## Where to go next

- Use `tempest_run()` for a scripted STORM report.
- Use `tempest_session()` or `run_app()` for interactive Co-STORM research.
- Add semantic retrieval with `ragnar` through `embed_fn` or a pre-built
  `ragnar_store`.
- Compile selected structured programs with `tempest_compile_programs()`, then
  pass the complete verified `TempestProgramSet` to STORM or Co-STORM.
- Resolve an accepted procedure with `tempest_governed_procedure_ref()`, bind
  it by stage in a ProgramSet, and pass its matching pinned `knowledge_view` to
  every governed run or session.
- Review provisional evidence with `tempest_promotion_bundle()` and
  `tempest_graft_plan()` before exercising Graft acceptance authority.
- Capture progress with `tempest_progress_collector()` and inspect events with
  `tempest_progress_event_data()`.
- Evaluate scripted and interactive workflows with `tempest_task()` and
  `tempest_costorm_task()`.

Start with a narrow topic, a small expert and question budget, and durable
output. Expand model roles, retrieval, parallelism, and evaluation only after
the first workflow behaves as intended.
